---
title: "Markdown"
author: "SK"
date: "July 23, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Code for reading in the dataset and/or processing the data
```{r}
activity<-read.csv("C:/Users/hp/documents/activity.csv")
activity$date<-as.Date(activity$date,"%Y-%m-%d")
```

## 2 Histogram of the total number of steps taken each day
```{r}
d1<-aggregate(x=activity[c("steps","interval")],FUN=sum,by=list(Group.date=activity$date),na.rm=TRUE)
hist(d1$steps,col="grey",border="black",xlab="Interval",main="Histogram-Total Steps per day")
```

## 3 Mean and median number of steps taken each day
```{r}
d1mean<-setNames(aggregate(x=activity[c("steps")],FUN=mean,by=list(activity$date),na.rm=TRUE),c("Date","Mean"))
d1median<-setNames(aggregate(x=activity[c("steps")],FUN=median,by=list(activity$date),na.rm=TRUE),c("Date","Median"))
d1stats<-merge(d1mean,d1median,by="Date")
```
```{r d1stats}
print(d1stats)
```

##4 Time series plot of the average number of steps taken
aggregate table 
```{r}
d2mean<- setNames(aggregate(x= activity[c("steps")],
FUN = mean,
by = list(activity$interval),na.rm=TRUE),c("Interval","MeanStep"))
```

Interval with Max Average Steps
```{r}
d2max<- subset(d2mean,d2mean$MeanStep==max(d2mean$MeanStep,na.rm=TRUE))
```

Plot Histogram
```{r}
plot(d2mean$Interval,d2mean$MeanStep,type="l",main="Steps per Interval",xlab="Interval",ylab="Step(Mean)",col="green",lwd=1)
abline(v=d2max$Interval,col ="blue",lwd=2)
legend(x="topright", c("MaxSteps Interval"),col=c("blue"),lwd=c(2,2,2))
```

## 5 The 5 mins interval that, on average, contains the maximum number of steps
```{r d2max}
d2max<-subset(d2mean,d2mean$MeanStep==max(d2mean$MeanStep,na.rm=TRUE))
print(d2max)
```

## 6 Code to describe and show a strategy for inputing missing data
create new variable
```{r}
d3<-activity
```
Replace NA values in column "Steps" with global average
```{r}
d3$steps[which(is.na(d3$steps))]<-mean(d3$steps,na.rm=TRUE)
```
```
heads(d3)
```

## 7 Histogram of the total number of steps taken each day after missing values are inputed
aggregate data for histogram
```{r}
d4<- aggregate(x=d3[c("steps")],
FUN=sum,
by=list(Group.date=activity$date),na.rm=TRUE)
```

# stats
```{r}
d4mean <- setNames(aggregate(x = d4[c("steps")],
FUN = mean,
by = list(d4$Group.date),na.rm=TRUE),c("Date","Mean"))
d4median<- setNames(aggregate(x=d4[c("steps")],
FUN = median,
by = list(d4$Group.date),na.rm=TRUE),c("Date","Median"))
d4stats<- merge(d4mean,d4median,by="Date")
```
```{r d4stats}
print(d4stats)
```
Plot Histogram
```{r}
hist(d4$steps,col="grey",border="black",xlab="SUM(steps)", main="Histogram Total Steps per day")
abline(v=mean(d4$steps),col="red",lwd=2)
abline(v=median(d4$steps),col="blue",lty=2,lwd=2)
legend(x="topright",c("mean","median"),col = c("red","blue"),lwd=c(2,2,2))
```

## 8 Panel plot comparing the average number of steps taken per 5-mins interval across weekdays and weekends
```{r}
d5<-d3
d5$day<- weekdays(d3$date)
d5$day<- ifelse(d5$day=="Saturday"|d5$day=="Sunday","weekend","weekday")
d5$day<-as.factor(d5$day)
```

aggregate data 
```{r}
d5sum<-setNames(aggregate(x=d5[c("steps")],FUN=sum,by=list(d5$interval,d5$day),na.rm=TRUE),c("Interval","day","SumStep"))
```

Subset data for graphics
```{r}
d5weekday<-subset(d5sum,d5sum$day=="weekday")
d5weekend<- subset(d5sum,d5sum$day=="weekend")
```

plot histogram
```{r}
par(mfrow=c(2,1))
plot(d5weekday$Interval,d5weekday$SumStep,main="Interval vs SumStep [weekdays]",type="l",col="green",xlab="Interval",ylab="SumStep")
plot(d5weekend$Interval,d5weekend$SumStep,main="Interval vs SumStep[weekend]",type="l",col="blue",xlab="Interval",ylab="SumStep")
```



